---
title: "TRex Datafile Calibration"
author: "Hannah Anderson"
date: "2023-11-17"
output: pdf_document
---

##Introduction

We used the automated tracking program TRex to track fish location during
behavioral recordings. When exporting datafiles, TRex automatically converts the
pixel values from the video to centimeters for analysis purposes. However,
for our analyses we require the original pixel values of the video in order to
cross reference with the position of the four corners of the experimental
chambers, the coordinates of which were located in the original pixel values.

Here, we first convert the X,Y coordinated from the initial exported TRex
datafiles to the original pixel coordinates. Then, we calibrate the coordinates
of both the new version of the datafiles and the collected corner coordinates
into centimeters based on the known length of the lower wall (as seen from the
video). These calibrated version of the datafiles will be used by both
data_conversion_movement.Rmd and data_conversion_social.Rmd.


##Prerequisites

This script will not function properly if there is insufficient available
memory on the computer. When insufficient memory is available to the script it
will typically hang or stall part way through processing datafiles and R will
need to be manually restarted. Occasionally R will instead crash. We advise
anyone using this script to minimize other memory usage by other programs on
their computer to avoid these problems.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

#filepath for importing and exporting the datafiles
importpath <- "E://Turbidity Flux Experiment with OMR/CSV_datafiles/"
exportpath <- "E://Turbidity Flux Experiment with OMR/calibrated_CSVs/"

#reading the file names into a list
list_csv <- list.files(strtrim(importpath, nchar(importpath)-1),
                       pattern = "*.csv")

#manually collected coordinates for each of the corners of the film area in pixels
corners <- read_csv("E://Turbidity Flux Experiment with OMR/film_frame_coordinates.csv")
#making the tankID column of 'corners' match the corresponding tank names from
#the fish datafiles
corners$tankID <- paste0("tank", corners$tankID)
```


## Calibration

```{r calibration}
for (k in 1:(length(list_csv))) {
  #importing current datafile
  fishfile <- read_csv(paste0(importpath, list_csv[k]), col_select = c(3:4, 6:8))
  #replacing inf values in datafile with NAs, which R handles better
  fishfile[sapply(fishfile, is.infinite)] <- NA
  
  #converting TRex's automated calibration values back into pixel values
  uncalib_fish <- fishfile
  #TRex equated 1920 pixels to 30 cm
  uncalib_fish$X <- fishfile$X*(1920/30)
  uncalib_fish$Y <- fishfile$Y*(1920/30)
  
  #extracting the tankID and recording number from the file name of the current
  #datafile
  file_str <- strsplit(list_csv[k], split = "_")
  #using the above to find the corresponding X,Y coordinate values from the
  #'corners' dataframe
  X_left <- corners$bottomLeftX[corners$tankID == file_str[[1]][1] &
                                corners$recording == parse_number(
                                                     file_str[[1]][2])]
  X_right <- corners$bottomRightX[corners$tankID == file_str[[1]][1] &
                                corners$recording == parse_number(
                                                     file_str[[1]][2])]
  Y_left <- corners$bottomLeftY[corners$tankID == file_str[[1]][1] &
                                corners$recording == parse_number(
                                                     file_str[[1]][2])]
  Y_right <- corners$bottomRightY[corners$tankID == file_str[[1]][1] &
                                corners$recording == parse_number(
                                                     file_str[[1]][2])]
  
  #using the 'corners' coordinates to calculate the distance (in pixels) of the
  #"bottom" wall of the film area
  knowndist <- sqrt((X_left - X_right)^2 + (Y_left - Y_right)^2)
  #the above distance is known to be 20 cm long
  calibration <- 20/knowndist
  
  #converting the pixel coordinates into centimeters using the calibration value
  calib_fish <- uncalib_fish
  calib_fish$X <- uncalib_fish$X*calibration
  calib_fish$Y <- uncalib_fish$Y*calibration
  
  #exporting the calibrated datafile
  write_csv(calib_fish, paste0(exportpath, "calibrated_", list_csv[k]))
}  #end of k loop


#also calibrating the corner coordinate values
calib_corners <- corners
for (c in 1:(length(corners$tankID))) {
  knowndist <- sqrt((corners$bottomLeftX[c] - corners$bottomRightX[c])^2 +
                     (corners$bottomLeftY[c] - corners$bottomRightY[c])^2)
  calibration <- 20/knowndist
  calib_corners[c,3:10] <- calib_corners[c,3:10]*calibration
}
#exporting the calibrated 'corners' datafile
#write_csv(calib_corners, "D://Turbidity Flux Experiment with OMR/calibrated_corner_coordinates.csv")
```