---
title: "TRex Datafile Calibration"
author: "Hannah Anderson"
date: "2023-11-17"
output: pdf_document
---

##Introduction

We used the automated tracking program TRex to track fish location during
behavioral recordings. When exporting datafiles, TRex automatically converts the
pixel values from the video to centimeters for analysis purposes. However,
for our analyses we require the original pixel values of the video in order to
cross reference with the position of the four corners of the experimental
chambers, the coordinates of which were located in the original pixel values.

Here, we first convert the X,Y coordinated from the initial exported TRex
datafiles to the original pixel coordinates. Then, we calibrate the coordinates
of both the new version of the datafiles and the collected corner coordinates
into centimeters based on the known length of the lower wall (as seen from the
video). These calibrated version of the datafiles will be used by both
data_conversion_movement.Rmd and data_conversion_social.Rmd.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setting the working directory
knitr::opts_knit$set(root.dir = "D://Turbidity Flux Experiment with OMR/CSV_datafiles")
library(tidyverse)
```

```{r file read, message = FALSE}
#reading the file names into a list
list_csv <- list.files(pattern = "*.csv")
#from the above list, creating a list of dataframes while excluding unused
#columns
masterlist <- lapply(list_csv, read_csv, col_select = c(3:4, 6:8))
#replacing inf with NA, which R handles better
inf2NA <- function(v) replace(v, is.infinite(v), NA)
fishlist <- lapply(masterlist, function(d) replace(d, TRUE, sapply(d, inf2NA)))
#manually collected coordinates for each of the corners of the film area in pixels
corners <- read_csv("D://Turbidity Flux Experiment with OMR/film_frame_coordinates.csv")
#filepath for exported files
filepath <- "D://Turbidity Flux Experiment with OMR/calibrated_CSVs/"
```


## Converting Back to Pixel Coordinates

```{r pixels}
#creating a copy of fishlist that can hold the converted values
fishpx <- fishlist

#converting TRex's automated calibration back into pixel values
for (k in 1:(length(list_csv))) {
  #TRex equated 1920 pixels to equal 30 cm
  fishpx[[k]]$X <- fishlist[[k]]$X*(1920/30)
  fishpx[[k]]$Y <- fishlist[[k]]$Y*(1920/30)
}

#creating a copy of fishpx that can hold the calibrated values
calib_list <- fishpx

#converting from pixel values into proper centimeter values
for (d in 1:(length(list_csv))) {
  
  #extracting the tankID and recording number from the file name of the current
  #datafile
  file_str <- strsplit(list_csv[d], split = "_")
  #using the above to point to the appropriate X,Y coordinate values in the
  #'corners' dataframe
  X_left <- corners$bottomLeftX[corners$tankID == file_str[[1]][1] &
                                corners$recording == parse_number(
                                                     file_str[[1]][2])]
  X_right <- corners$bottomRightX[corners$tankID == file_str[[1]][1] &
                                corners$recording == parse_number(
                                                     file_str[[1]][2])]
  Y_left <- corners$bottomLeftY[corners$tankID == file_str[[1]][1] &
                                corners$recording == parse_number(
                                                     file_str[[1]][2])]
  Y_right <- corners$bottomRightY[corners$tankID == file_str[[1]][1] &
                                corners$recording == parse_number(
                                                     file_str[[1]][2])]
  
  #calculating the distance (in pixels) of the "bottom" wall of the film area
  knowndist <- sqrt((X_left - X_right)^2 + (Y_left - Y_right)^2)
  #the above distance is known to be 20 cm long
  calibration <- 20/knowndist
  
  #converting the pixel coordinates into centimeters using the calibration value
  calib_list[[d]]$X <- fishpx[[d]]$X*calibration
  calib_list[[d]]$Y <- fishpx[[d]]$Y*calibration
  #and then I need to export the datafiles
  write_csv(calib_list[[d]], paste0(filepath, "calibrated_", list_csv[[d]]))
}
```