---
title: "Flux Turbidity Data Conversion for Proximity Tests"
author: "Hannah Anderson"
date: "2023-11-29"
output: pdf_document
---

## Introduction

Data for this experiment was captured using vertically mounted cameras observing
the movement of zebrafish in shallow waters. The videos were then processed
using the automated tracking program TRex, which returns individual files for
each tracked fish for each video. Here, we are taking these individual fish
files and 1) converting the frame-by-frame data into summary statistics for each
video and 2) combining the data from the 900+ files into a singular dataset.
This dataset is then exported to be used in statistical analyses.

In this script specifically we are calculating values relating to proximity of
fish to other fish and the external walls of the experimental chamber as well as
polarity values for the fish shoals. See data_conversion_movement.Rmd for the
calculations of general movement patterns. If using one of these scripts after
the other make sure to clear the working environment as many variable names are
shared between scripts.


### Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setting the working directory
knitr::opts_knit$set(root.dir = "D://Turbidity Flux Experiment with OMR/calibrated_CSVs")
#loading packages
library(tidyverse)
library(data.table)
library(zoo)
library(sf)
library(circular)
#the number of fish in a group
fish = 4
```


### Importing Files

```{r file read in, message = FALSE}
#reading the file names into a list
list_csv <- list.files(pattern = "*.csv")
#from the file list, creating a list of dataframes
masterlist <- lapply(list_csv, read_csv)
#importing a list of coordinates for the corners of the film area for all videos
corners <- read_csv("D://Turbidity Flux Experiment with OMR/calibrated_corner_coordinates.csv")
###In the other file here is where I create a blank dataframe to contain the final data
```


### Processing Files

```{r loops, include = FALSE, message = FALSE}
#enumerating the file number
i = 1

#probably should put a comment here
for (k in 1:(length(list_csv)/fish)) {  #across the whole file list
  
  #combining the files from all fish per video to be analyzed collectively
  shoal <- masterlist[i:(i+fish-1)]
  
  
  ##standardizing the length of each fish's file in a video so all files start
  ##at frame 0 and end at the last frame of the longest file
  
  #determining the last frame of a video by taking the largest frame value
  #from all of the fish datafiles
  maxframe <- max(unlist(lapply(shoal, function(m) m[,"frame"])))
  #creating a dataframe filled with NAs to pad the missing frames from each file
  dfNA <- data.frame(frame = 0:maxframe, fish_detected = NA, angle = NA,
                     X = NA, Y = NA)
  
  
  
  for (z in 1:fish) {  #across one video of files
    
    #finding the starting frame of the current datafile
    minfish <- min(shoal[[z]][,"frame"])
    
    #adding NA rows from dfNA to the start of a file when it starts on a frame
    #later than 0
    if (minfish > 0) {  #across one fish file within a video
      shoal[[z]] <- rbind(dfNA[1:(minfish),], shoal[[z]])
    }
    
    #finding the frame the current datafile ends on
    maxfish <- max(shoal[[z]][,"frame"])
    
    #adding NA rows from dfNA to the end of a file when it ends earlier than the
    #longest file from the current video
    if(maxfish < maxframe) {  #across one fish file within a video
      shoal[[z]] <- rbind(shoal[[z]], dfNA[(maxfish+2):(maxframe+1),])
    }
  } #z loop end
  
  
  
  #creating an empty dataframe to hold the neighbor distance values
  distances <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 6))
  colnames(distances) <- c(1:6)
  #creating an empty dataframe to hold the polarity and shoal area values
  shoal_angle <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 2))
  colnames(shoal_angle) <- c("shoal_area", "polarity")
  
  #calculating nearest and farthest neighbor, polarity and shoal area
  for (a in 1:(maxframe+1)) {  #across one video of files
    
    #pulling the X,Y coordinates from each of the fish files in a video
    coor <- data.frame(matrix(nrow = fish, ncol = 2))
    colnames(coor) <- c("X", "Y")
    for (p in 1:fish) {
      coor$X[p] <- shoal[[p]][a,"X"]
      coor$Y[p] <- shoal[[p]][a,"Y"]
    }
    
    #calculating the distances between every fish
    neighbors <- as.vector(dist(coor))
    
    #adding the distance values into the distance dataframe
    for (d in 1:(length(neighbors))) {
      distance <- neighbors[d]
    }
    
    #calculating shoal areas, but only if all fish in the group are detected
    #for that frame
    if (anyNA(as.vector(unlist(coor))) == FALSE) {
      #sf needs the first and last coordinates of a polygon to be the same to
      #finish drawing it
      coorpoly <- rbind(coor, coor[1,])
      #changing the classes of the columns back to numeric from list
      coorpoly$X <- unlist(coorpoly$X)
      coorpoly$Y <- unlist(coorpoly$Y)
      #converting the coordinates to a polygon and adding to the polarity and
      #shoal area dataframe
      shoal_angle$shoal_area[a] <- st_area(st_polygon(
                                           x = list(cbind(
                                                    coorpoly$X,coorpoly$Y))))
    }
    
    
    #calculating polarity, which is the variance of fish angles- since this
    #data is inherently circular we're using circular variance
    
    #converting the data into a "circular" object for the package 'circular'
    circ <- suppressWarnings(as.circular(unlist(lapply(shoal, function(v)
                                         v[a,"angle"])), type = "angles",
                                         units = "radians", rotation = "clock",
                                         zero = 0))
    shoal.angle$polarity[a] <- var.circular(circ)
    
    
    #calculating the distance between the fish and the nearest wall
    
    #extracting the tankID and recording number from the file name of the
    #current datafile
    file_str <- strsplit(list_csv[i], split = "_")
    #creating a dataframe to hold the coordinates for each wall of the film
    #area
    walls <- data.frame(matrix(nrow = 4, ncol = 4))
    colnames(walls) <- c("Ax", "Ay", "Bx", "By")
    #bottom wall
    walls[1,] <- corners[corners$tankID == file_str[[1]][2] &
                                corners$recording == parse_number(
                                                     file_str[[1]][3]
                              ),3:6]
    #left wall
    walls[2,] <- corners[corners$tankID == file_str[[1]][2] &
                                corners$recording == parse_number(
                                                     file_str[[1]][3]
                              ),c(3:4,9:10)]
    #right wall
    walls[3,] <- corners[corners$tankID == file_str[[1]][2] &
                                corners$recording == parse_number(
                                                     file_str[[1]][3]
                              ),c(5:6,7:8)]
    #top wall
    walls[4,] <- corners[corners$tankID == file_str[[1]][2] &
                                corners$recording == parse_number(
                                                     file_str[[1]][3]
                              ),c(9:10,7:8)]
    #empty dataframe to hold the wall proximity values
    wall_dist <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 4))
    
    
    #calculate the distance for each wall, then take the smallest value
    #I realize now this entire file is probably not the most efficient way to be
    #doing things but it's way too late to fix that now
    for (f in (1:fish)) {
      tempdist <- c(NA, NA, NA, NA)
      for (w in 1:4) {
        #wall coordinates 2 - fish coordinates
        v1 <- as.numeric(walls[w,1:2] - walls[w,3:4])
        #wall coordinates 1 - wall coordinates 2
        v2 <- as.numeric(shoal[[f]][a,4:5] - walls[w,1:2])
        m <- cbind(v1,v2)
        tempdist[w] <- abs(det(m))/sqrt(sum(v1^2))
      }
      wall_dist[a,f] <- min(tempdist)
    }
      
  } #a loop end
  

  
  
  ###down here will go the noise corrections
}
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
