---
title: "Flux Turbidity Data Conversion for Proximity Tests"
author: "Hannah Anderson"
date: "2023-11-29"
output: pdf_document
---

## Introduction

Data for this experiment was captured using vertically mounted cameras observing
the movement of zebrafish in shallow waters. The videos were then processed
using the automated tracking program TRex, which returns individual files for
each tracked fish for each video. Here, we are taking these individual fish
files and 1) converting the frame-by-frame data into summary statistics for each
video and 2) combining the data from the 900+ files into a singular dataset.
This dataset is then exported to be used in statistical analyses.

In this script specifically we are calculating values relating to proximity of
fish to other fish and the external walls of the experimental chamber as well as
polarity values for the fish shoals. See data_conversion_movement.Rmd for the
calculations of general movement patterns. If using one of these scripts after
the other make sure to clear the working environment as many variable names are
shared between scripts.

NOTE: MAKE SURE TO CHANGE THE "experiment" VARIABLE IN SETUP TO THE APPROPRIATE
VALUE OR THE CODE WILL EXCLUDE TWO VIDEOS FROM SHOAL AREA CALCULATIONS.


### Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setting the working directory
knitr::opts_knit$set(root.dir = "D://Turbidity Flux Experiment with OMR/calibrated_CSVs")
#loading packages
library(tidyverse)
library(data.table)
library(zoo)
library(sf)
library(circular)
#the number of fish in a group
fish = 4

#if processing the clay experiment, make the string "clay", if processing the
#dye experiment, make the string "dye"
experiment <- "clay"
#the tank IDs in each treatment for the clay experiment
clayclear <- c("tank1C", "tank5B", "tank3C", "tank8A", "tank9A", "tank4C",
               "tank6A", "tank6B", "tank2C", "tank2B")
claystable <- c("tank1B", "tank9B", "tank8B", "tank3B", "tank9C", "tank7A",
                "tank3A", "tank6C", "tank1A", "tank10A")
clayflux <- c("tank10C", "tank8C", "tank10B", "tank5C", "tank2A", "tank7C",
             "tank5A", "tank4B", "tank4A", "tank7B")
#the tank IDs in each treatment for the dye experiment
dyeclear <- c("tank2A", "tank8C", "tank4C", "tank1C", "tank8B", "tank9C",
              "tank3A", "tank4A", "tank5B", "tank1B")
dyestable <- c("tank6A", "tank3B", "tank6C", "tank4B", "tank7B", "tank6B",
               "tank7C", "tank7A", "tank10C", "tank9A")
dyeflux <- c("tank9B", "tank3C", "tank10A", "tank8A", "tank1A", "tank5C",
             "tank10B", "tank2C", "tank2B")
```


### Importing Files

```{r file read in, message = FALSE}
#reading the file names into a list
list_csv <- list.files(pattern = "*.csv")
#from the file list, creating a list of dataframes
masterlist <- lapply(list_csv, read_csv)
#importing a list of coordinates for the corners of the film area for all videos
corners <- read_csv("D://Turbidity Flux Experiment with OMR/calibrated_corner_coordinates.csv")
#creating a dataframe to hold the final data
finalframe <- data.frame(matrix(ncol = 11, nrow = (length(list_csv)/fish)))
colnames(finalframe) <- c("tankID", "treatment", "recording", "polarity",
                          "shoal_area", "shoal_area_sd", "nearest_neighbor",
                          "farthest_neighbor", "mean_neighbor",
                          "wall_proximity", "space_use")
```


### Processing Files

```{r loops, include = FALSE, message = FALSE}
#enumerating the file number
i = 1

#probably should put a comment here
for (k in 1:(length(list_csv)/fish)) {  #across the whole file list
  
  #combining the files from all fish per video to be analyzed collectively
  shoal <- masterlist[i:(i+fish-1)]
  
  #empty dataframe to hold the wall proximity values
  wall_dist <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 4))
  colnames(wall_dist) <- c(1:4)
  
  ##standardizing the length of each fish's file in a video so all files start
  ##at frame 0 and end at the last frame of the longest file
  
  #determining the last frame of a video by taking the largest frame value
  #from all of the fish datafiles
  maxframe <- max(unlist(lapply(shoal, function(m) m[,"frame"])))
  #creating a dataframe filled with NAs to pad the missing frames from each file
  dfNA <- data.frame(frame = 0:maxframe, fish_detected = NA, angle = NA,
                     X = NA, Y = NA)
  
  
  
  for (z in 1:fish) {  #across one video of files
    
    #finding the starting frame of the current datafile
    minfish <- min(shoal[[z]][,"frame"])
    
    #adding NA rows from dfNA to the start of a file when it starts on a frame
    #later than 0
    if (minfish > 0) {  #across one fish file within a video
      shoal[[z]] <- rbind(dfNA[1:(minfish),], shoal[[z]])
    }
    
    #finding the frame the current datafile ends on
    maxfish <- max(shoal[[z]][,"frame"])
    
    #adding NA rows from dfNA to the end of a file when it ends earlier than the
    #longest file from the current video
    if(maxfish < maxframe) {  #across one fish file within a video
      shoal[[z]] <- rbind(shoal[[z]], dfNA[(maxfish+2):(maxframe+1),])
    }
  } #z loop end
  
  
  
  #creating an empty dataframe to hold the neighbor distance values
  distances <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 6))
  colnames(distances) <- c(1:6)
  #creating an empty dataframe to hold the polarity and shoal area values
  shoal_angle <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 2))
  colnames(shoal_angle) <- c("shoal_area", "polarity")
  
  #extracting the tankID and recording number from the file name of the
  #current datafile
  file_str <- strsplit(list_csv[i], split = "_")
  
  #creating a dataframe to hold the coordinates for each wall of the film area
  #for use in calculating fish proximity to the nearest wall
  walls <- data.frame(matrix(nrow = 4, ncol = 4))
  colnames(walls) <- c("Ax", "Ay", "Bx", "By")
  #bottom wall
  walls[1,] <- corners[corners$tankID == file_str[[1]][2] &
                       corners$recording == parse_number(
                                            file_str[[1]][3]
                                            ),3:6]
  #left wall
  walls[2,] <- corners[corners$tankID == file_str[[1]][2] &
                       corners$recording == parse_number(
                                            file_str[[1]][3]
                                            ),c(3:4,9:10)]
  #right wall
  walls[3,] <- corners[corners$tankID == file_str[[1]][2] &
                       corners$recording == parse_number(
                                            file_str[[1]][3]
                                            ),c(5:6,7:8)]
  #top wall
  walls[4,] <- corners[corners$tankID == file_str[[1]][2] &
                       corners$recording == parse_number(
                                            file_str[[1]][3]
                                            ),c(9:10,7:8)]
  
  #creating an empty dataframe to hold the coordinates for the corners of the
  #film area for use in calculating shoal space use
  film_area <- data.frame(matrix(nrow = 2, ncol = 4))
  colnames(film_area) <- c("X", "Y")
  #adding the film area coordinates to the dataframe, repeating the first set
  #of coordinates because sf needs the first and last coordinates of a polygon
  #to be the same to "complete" the polygon
  film_area[1,] <- corners[corners$tankID == file_str[[1]][2] &
                       corners$recording == parse_number(
                                            file_str[[1]][3]
                                            ),c(3:4)]
  film_area[2,] <- corners[corners$tankID == file_str[[1]][2] &
                       corners$recording == parse_number(
                                            file_str[[1]][3]
                                            ),c(5:6)]
  film_area[3,] <- corners[corners$tankID == file_str[[1]][2] &
                       corners$recording == parse_number(
                                            file_str[[1]][3]
                                            ),c(7:8)]
  film_area[4,] <- corners[corners$tankID == file_str[[1]][2] &
                       corners$recording == parse_number(
                                            file_str[[1]][3]
                                            ),c(9:10)]
  film_area[5,] <- corners[corners$tankID == file_str[[1]][2] &
                       corners$recording == parse_number(
                                            file_str[[1]][3]
                                            ),c(3:4)]
  #creating an sf polygon from the film area coordinates
  film_poly <- st_polygon(x = list(as.matrix(film_area)))
  #creating a grid from the film area polygon that's 20 cells wide by 24 cells
  #high to match the 20cm x 24cm filming area
  film_grid <- st_make_grid(film_poly, n = c(20,24), what = "polygons",
                            square = TRUE)
  #adding IDs to each grid cell
  film_grid_ID <- st_sf(film_grid) %>%
          mutate(grid_id = 1:length(lengths(film_grid)))
  
  #calculating the number of frames each grid cell is occupied by each fish
  #during the video
  
  #converting the fish coordinates to sf objects
  fish0_coor <- shoal[[1]] %>%
                st_as_sf(coords = c("X", "Y"), remove = FALSE,
                          na.fail = FALSE)
  fish1_coor <- shoal[[2]] %>%
                st_as_sf(coords = c("X", "Y"), remove = FALSE,
                          na.fail = FALSE)
  fish2_coor <- shoal[[3]] %>%
                st_as_sf(coords = c("X", "Y"), remove = FALSE,
                          na.fail = FALSE)
  fish3_coor <- shoal[[4]] %>%
                st_as_sf(coords = c("X", "Y"), remove = FALSE,
                          na.fail = FALSE)
  #detecting the number of frames each fish in each grid cell by counting the
  #number of intersections between the fish location datapoints and the
  #polygons of each grid cell
  film_grid_ID$fish0 <- lengths(st_intersects(film_grid_ID, fish0_coor))
  film_grid_ID$fish1 <- lengths(st_intersects(film_grid_ID, fish1_coor))
  film_grid_ID$fish2 <- lengths(st_intersects(film_grid_ID, fish2_coor))
  film_grid_ID$fish3 <- lengths(st_intersects(film_grid_ID, fish3_coor))
  
  #the below commented-out code can be used to produce plots to test that the
  #above detection code worked properly, with the red line overlaying each plot
  #representing the film area and the color of the grid cells representing how
  #many frames a fish occupied that space
  #library(ggpubr)
  #testplot1 <- ggplot(film_grid_ID) +
              #geom_sf(aes(fill = fish0)) +
              #scale_fill_distiller(palette = "PiYG") +
              #geom_sf(data = st_zm(film_poly), fill = NA, color = "red")
  #testplot2 <- ggplot(film_grid_ID) +
              #geom_sf(aes(fill = fish1)) +
              #scale_fill_distiller(palette = "PiYG") +
              #geom_sf(data = st_zm(film_poly), fill = NA, color = "red")
  #testplot3 <- ggplot(film_grid_ID) +
              #geom_sf(aes(fill = fish2)) +
              #scale_fill_distiller(palette = "PiYG") +
              #geom_sf(data = st_zm(film_poly), fill = NA, color = "red")
  #testplot4 <- ggplot(film_grid_ID) +
              #geom_sf(aes(fill = fish3)) +
              #scale_fill_distiller(palette = "PiYG") +
              #geom_sf(data = st_zm(film_poly), fill = NA, color = "red")
  #show <- ggarrange(testplot1, testplot2, testplot3, testplot4,
                    #ncol = 2, nrow = 2)
  
  #adding the intersections for all fish together for analysis as a group
  film_grid_ID$total <- NA
  for (x in 1:(length(film_grid_ID$grid_id))) {
    film_grid_ID$total[x] <- sum(film_grid_ID$fish0[x], film_grid_ID$fish1[x],
                                 film_grid_ID$fish2[x], film_grid_ID$fish3[x])
  }
  
  #testplotx <- ggplot(film_grid_ID) +
              #geom_sf(aes(fill = total)) +
              #scale_fill_distiller(palette = "PiYG") +
              #geom_sf(data = st_zm(film_poly), fill = NA, color = "red")
  
  #binary yes/no defining if a cell has been used based on a threshold of
  #'thresh' total intersections per cell
  film_grid_ID$used <- NA
  thresh <- 30
  for (t in 1:(length(film_grid_ID$grid_id))) {
    if (film_grid_ID$total[t] >= thresh) {
      film_grid_ID$used[t] <- 1
    } else
      film_grid_ID$used[t] <- 0
  }
  
  #length(which(film_grid_ID$used == 1))
  #length(which(film_grid_ID$used == 0))
  
  #useplot <- ggplot(film_grid_ID) +
    #geom_sf(aes(fill = used)) +
    #scale_fill_distiller(palette = "PiYG") +
    #geom_sf(data = st_zm(film_poly), fill = NA, color = "red")
  #library(ggpubr)
  #show <- ggarrange(testplotx, useplot, ncol = 2, nrow = 1)
  
  #calculating nearest and farthest neighbor, polarity and shoal area
  for (a in 1:(maxframe+1)) {  #across one video of files
    
    #pulling the X,Y coordinates from each of the fish files in a video
    coor <- data.frame(matrix(nrow = fish, ncol = 2))
    colnames(coor) <- c("X", "Y")
    for (p in 1:fish) {
      coor$X[p] <- shoal[[p]][a,"X"]
      coor$Y[p] <- shoal[[p]][a,"Y"]
    }
    
    #calculating the distances between every fish
    neighbors <- as.vector(dist(coor))
      
    #adding the distance values into the distance dataframe
    for (d in 1:(length(neighbors))) {
      distances[a,d] <- neighbors[d]
    }
    
  
    
    #calculating shoal areas, but only if all fish in the group are detected
    #for that frame
    if (anyNA(as.vector(unlist(coor))) == FALSE) {
      #sf needs the first and last coordinates of a polygon to be the same to
      #finish drawing it
      coorpoly <- rbind(coor, coor[1,])
      #changing the classes of the columns back to numeric from list
      coorpoly$X <- unlist(coorpoly$X)
      coorpoly$Y <- unlist(coorpoly$Y)
      #converting the coordinates to a polygon and adding to the polarity and
      #shoal area dataframe
      shoal_angle$shoal_area[a] <- st_area(st_polygon(
                                           x = list(cbind(
                                                    coorpoly$X,coorpoly$Y))))
    #in the clay experiment, tank 3A lost one fish, giving it only three fish
    #but files for four fish. The below code detects if the active file is one
    #of the videos that contains only three fish and calculates shoal area if
    #all three are present
    } else if (experiment == "clay" & file_str[[1]][2] == "tank3A" &
               file_str[[1]][3] != "record1" & file_str[[1]][3] != "record2") {
      #due to there being four fish files but only three fish it's expected for
      #there to be one NA in the X column and one in Y (one for each of the X,Y
      #coordinates). This only calculates the shoal area if there are the
      #expected NA values
      if (sum(is.na(as.vector(coor$X))) == 1 &
        sum(is.na(as.vector(coor$Y))) == 1) {
        coor$X <- unlist(coor$X)
        coor$Y <- unlist(coor$Y)
        #removing the row with NAs so st_area() can process the polygon
        coor <- na.omit(coor)
        coorpoly <- rbind(coor, coor[1,])
        shoal_angle$shoal_area[a] <- st_area(st_polygon(
                                               x = list(cbind(
                                                        coorpoly$X,coorpoly$Y))))
      }
    #the dye experiment also lost one fish before the third recording, though
    #this time from tank 1A. The below code does the same as the above code does
    #for the clay experiment's tank 3A videos
    } else if (experiment == "dye" & file_str[[1]][2] == "tank1A" &
               file_str[[1]][3] != "record1" & file_str[[1]][3] != "record2") {
      if (sum(is.na(as.vector(coor$X))) == 1 &
          sum(is.na(as.vector(coor$Y))) == 1) {
        coor$X <- unlist(coor$X)
        coor$Y <- unlist(coor$Y)
         coor <- na.omit(coor)
        coorpoly <- rbind(coor, coor[1,])
        shoal_angle$shoal_area[a] <- st_area(st_polygon(
                                              x = list(cbind(
                                                       coorpoly$X,coorpoly$Y))))
      }
    }
    
    #calculating polarity, which is the variance of fish angles- since this
    #data is inherently circular we're using circular variance
    ##only calculates polarity if all fish are detected
    
    #converting the data into a "circular" object for the package 'circular'
    circ <- suppressWarnings(as.circular(unlist(lapply(shoal, function(v)
                                         v[a,"angle"])), type = "angles",
                                         units = "radians", rotation = "clock",
                                         zero = 0))
    #the same as with shoal area, the below code determines if the current
    #video is one of the ones containing three fish instead of four and, if it
    #is, allows for the polarity to be calculated with only three fish
    if (experiment == "clay" & file_str[[1]][2] == "tank3A" &
               file_str[[1]][3] != "record1" & file_str[[1]][3] != "record2") {
      if (sum(is.na(circ)) == 1) {
        shoal_angle$polarity[a] <- var.circular(circ, na.rm = TRUE)
      } else shoal_angle$polarity[a] <- NA
    } else if (experiment == "dye" & file_str[[1]][2] == "tank1A" &
               file_str[[1]][3] != "record1" & file_str[[1]][3] != "record2") {
      if (sum(is.na(circ)) == 1) {
        shoal_angle$polarity[a] <- var.circular(circ, na.rm = TRUE)
      } else shoal_angle$polarity[a] <- NA
    } else {
          shoal_angle$polarity[a] <- var.circular(circ)
    }
    
    
    #calculating the distance between the fish and the nearest wall
    
    #calculate the distance for each wall, then take the smallest value
    for (f in (1:fish)) {
      tempdist <- c(NA, NA, NA, NA)
      for (w in 1:4) {
        #wall coordinates 2 - fish coordinates
        v1 <- as.numeric(walls[w,1:2] - walls[w,3:4])
        #wall coordinates 1 - wall coordinates 2
        v2 <- as.numeric(shoal[[f]][a,4:5] - walls[w,1:2])
        m <- cbind(v1,v2)
        tempdist[w] <- abs(det(m))/sqrt(sum(v1^2))
      }
      wall_dist[a,f] <- min(tempdist)
    }
  } #a loop end
  
  
  
  #correcting for the noise in TRex's detection of fish locations using rolling
  #averages
  
  #creating an empty dataframe to hold the rolling averages for neighbor
  #distances
  roll_neigh <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 6))
  colnames(roll_neigh) <- c(1:6)
  #calculating the rolling averages for neighbor distances
  for (r in 1:6) {
    roll_neigh[r] <- rollapply(distances[,r], width = 7, FUN = function(x)
                             mean(x, na.rm = TRUE), partial = TRUE, fill = NA)
  }
  
  #creating an empty dataframe to hold the rolling averages for polarity and
  #shoal area
  roll_polarea <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 2))
  colnames(roll_polarea) <- c("shoal_area", "polarity")
  #calculating the rolling average for shoal area
  roll_polarea$shoal_area <- rollapply(shoal_angle$shoal_area, width = 7,
                                       FUN = function(x) mean(x, na.rm = TRUE),
                                       partial = TRUE, fill = NA)
  #calculating the rolling average for polarity
  roll_polarea$polarity <- rollapply(shoal_angle$polarity, width = 7,
                                       FUN = function(x) mean(x, na.rm = TRUE),
                                       partial = TRUE, fill = NA)
  
  #creating an empty dataframe to hold the rolling averages for wall proximity
  roll_wall <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 4))
  colnames(roll_wall) <- c(1:4)
  #calculating the rolling mean for wall proximity
  for (r in 1:4) {
    roll_wall[r] <- rollapply(wall_dist[,r], width = 7, FUN = function(x)
                             mean(x, na.rm = TRUE), partial = TRUE, fill = NA)
  }
  
  
  #exporting the data
  
  #using the filename information to label the tank ID and recording number for
  #each row of the datafile
  finalframe$tankID[k] <- file_str[[1]][2]
  finalframe$recording[k] <- file_str[[1]][3]
  #using the tankID and the experiment name to correctly assign treatment
  if (experiment == "clay") {
    if (finalframe$tankID[k] %in% clayclear == "TRUE") {
      finalframe$treatment[k] <- "control"
    } else if (finalframe$tankID[k] %in% claystable == "TRUE") {
      finalframe$treatment[k] <- "stable"
    } else if (finalframe$tankID[k] %in% clayflux == "TRUE") {
      finalframe$treatment[k] <- "flux"
    } else {
      finalframe$treatment[k] <- "clay_error"
    }
  } else if (experiment == "dye") {
    if (finalframe$tankID[k] %in% dyeclear == "TRUE") {
      finalframe$treatment[k] <- "control"
    } else if (finalframe$tankID[k] %in% dyestable == "TRUE") {
      finalframe$treatment[k] <- "stable"
    } else if (finalframe$tankID[k] %in% dyeflux == "TRUE") {
      finalframe$treatment[k] <- "flux"
    } else {
      finalframe$treatment[k] <- "dye_error"
    }
  }
  
  #adding the data calculated above to the final dataframe
  finalframe$nearest_neighbor[k] <- min(unlist(roll_neigh, use.names = FALSE),
                                     na.rm = TRUE)
  finalframe$farthest_neighbor[k] <- max(unlist(roll_neigh, use.names = FALSE),
                                      na.rm = TRUE)
  finalframe$mean_neighbor[k] <- mean(unlist(roll_neigh, use.names = FALSE),
                                      na.rm = TRUE)
  finalframe$shoal_area[k] <- mean(roll_polarea$shoal_area, na.rm = TRUE)
  finalframe$shoal_area_sd[k] <- stats::sd(roll_polarea$shoal_area, na.rm = TRUE)
  finalframe$wall_proximity[k] <- mean(unlist(roll_wall, use.names = FALSE),
                                    na.rm = TRUE)
  #high circular variance equals low polarity, so subtracting the variance from
  #one to make the values more intuitive
  finalframe$polarity[k] <- 1 - (mean(roll_polarea$polarity, na.rm = TRUE))
  #taking the proportion of tank area grid cells used to the total number of
  #cells (used or unused) as a measure of total space use by the shoal
  finalframe$space_use[k] <- (length(which(film_grid_ID$used == 1)))/
                             (length(which(film_grid_ID$used == 0)) +
                              length(which(film_grid_ID$used == 1)))
  
  i = i+fish
} #k loop end

write_csv(finalframe, "D://Turbidity Flux Experiment with OMR/6day_clay_proximity_data.csv")
```
