---
title: "Statistical Analyses of TRex Proximity Data"
author: "Hannah Anderson"
date: "2024-02-26"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Introduction

WIP: We ran two models on each of our measures. The first is a model
investigating the effects across all three treatments specifically during
recording 3 (the "direct comparison" models). This recording occurs when both
the stable and flux treatments are at the same visibility level but also after
two previous fluctuations had occurred during the flux condition (during the
previous day). These models include an interaction effect between the experiment
(clay or dye) and treatment as well as a random effect of tank ID.

The second set of models includes all of the recordings from all experiments and
includes an interaction between the experiment (clay or dye), treatment and the
visibility of the flux condition/time of day as well as a random effect of 
tank ID.


## Setup

```{r setup, include=FALSE}
setwd("C://Users/hande/Documents/R/fluctuating-turbidity2")
library(tidyverse)
library(glmmTMB)
library(emmeans)
library(car)
library(DHARMa)

#importing the data
clay_data <- read_csv("6day_clay_proximity_data.csv")
dye_data <- read_csv("6day_dye_proximity_data.csv")
#adding a column to identify the experiment
clay_data$experiment <- "clay"
dye_data$experiment <- "dye"

#changing tank IDs to shoal IDs to ensure IDs are different between experiments
clay_data <- clay_data %>%
             rename_at("tankID", ~"shoalID")
clay_data$shoalID <- gsub("tank", "clay", clay_data$shoalID)
dye_data <- dye_data %>%
            rename_at("tankID", ~"shoalID")
dye_data$shoalID <- gsub("tank", "dye", dye_data$shoalID)

#combining the clay and dye data together into a single data frame
comp_data <- rbind(clay_data, dye_data)
```


## Direct Comparison Models

```{r direct}
#sub-setting the data to only include recording 3
direct_data <- comp_data %>%
               filter(recording == "record3")

#mean shoal area
d_shoal <- glmmTMB(shoal_area ~ treatment*experiment,
                   data = direct_data, gaussian())
#mean neighbor distance
##model convergence
d_neigh <- update(d_shoal, mean_neighbor ~ .)
#mean proximity to wall
d_wall <- update(d_shoal, wall_proximity ~ .)
#polarity
d_polar <- update(d_shoal, polarity ~ .)
#ratio of space use
d_space <- update(d_shoal, space_use ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = d_shoal))
plot(simulateResiduals(fittedModel = d_neigh))
plot(simulateResiduals(fittedModel = d_wall))
plot(simulateResiduals(fittedModel = d_polar)) #outlier test positive
plot(simulateResiduals(fittedModel = d_space)) #assumptions not met

#transforming the data to meet model assumptions
trans_data <- comp_data
trans_data$space_use <- (trans_data$space_use)^2

#the transformed model
trans_space <- update(d_space, . ~ ., data = trans_data)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = trans_space))

#analyzing the results
Anova(d_shoal)
Anova(d_neigh)
Anova(d_wall)
Anova(d_polar)
#Anova(d_space)

#post hoc pairwise comparisons of the significant treatment results
d_shoal %>% emmeans(pairwise ~ treatment)
d_wall %>% emmeans(pairwise ~ treatment)
d_wall %>% emmeans(pairwise ~ treatment:experiment)
```


## Full Comparison Models

```{r full}
#adding a column to define whether the recording is in the morning (when the
#flux visibility is the same as the stable visibility) or in the afternoon (when
#flux visibility is better than the stable visibility)
comp_data$time <- comp_data$recording
comp_data$time <- comp_data$time %>%
                  fct_collapse(morning = c("record1", "record3"),
                               afternoon = c("record2", "record4"))

####still need to average!

#mean shoal area
m_shoal <- glmmTMB(shoal_area ~ treatment*experiment*time + (1|shoalID),
                   data = comp_data, family = gaussian())
#mean neighbor distance
m_neigh <- update(m_shoal, mean_neighbor ~ .)
#mean proximity to wall
m_wall <- update(m_shoal, wall_proximity ~ .)
#polarity
polar <- update(m_shoal, polarity ~ ., family = beta_family(link = "logit"))
#ratio of space use
space <- update(polar, space_use ~ .)

#checking model assumptions
plot(simulateResiduals(fittedModel = m_shoal)) #assumptions borderline
plot(simulateResiduals(fittedModel = m_neigh))
plot(simulateResiduals(fittedModel = m_wall))
plot(simulateResiduals(fittedModel = polar)) #assumptions not met
plot(simulateResiduals(fittedModel = space)) #assumptions not met

#analyzing the results
#Anova(m_shoal)
Anova(m_neigh)
Anova(m_wall)
#Anova(polar)
#Anova(space)
```